---
title: "Summary statistics on the surface texture parameters for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_3b_STA-stats.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
This script computes standard descriptive statistics for each group.  
The groups are based on:

* Chert type (A or B)  
* Bamboo species (Bambusa blumeana or Schizostachum lima)  
* Strokes (0 or 2000)

It computes the following statistics:  

* sample size (n = `length`)  
* smallest value (`min`)  
* largest value (`max`)
* mean  
* median  
* standard deviation (`sd`)

```{r}
dir_in <- "analysis/derived_data"
dir_stats <- "analysis/stats/"
```

Raw data must be located in "`r paste0("./", dir_in)`".  
Summary statistics will be saved in "`r paste0("./", dir_stats)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(doBy)
library(grateful)
library(knitr)
library(R.utils)
library(rmarkdown)
library(tidyverse)
library(writexl)
```


---


# Read in data
```{r}
STA <- list.files(dir_in, pattern = "STA\\.Rbin$", full.names = TRUE) %>% 
       loadObject()
str(STA)
head(STA)
```


---


# Summary statistics

## Create function to compute the statistics at once
```{r}
nminmaxmeanmedsd <- function(x){
    y <- x[!is.na(x)]     # Exclude NAs
    n_test <- length(y)   # Sample size (n)
    min_test <- min(y)    # Minimum
    max_test <- max(y)    # Maximum
    mean_test <- mean(y)  # Mean
    med_test <- median(y) # Median
    sd_test <- sd(y)      # Standard deviation
    out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test) # Concatenate
    names(out) <- c("n", "min", "max", "mean", "median", "sd")         # Name values
    return(out)                                                        # Object to return
}
```


## Compute summary statistics
```{r}
# Exclude Chert_tool and Location columns from data because they are numeric yet not relevant for stats
STA_sel <- select(STA, !c(Chert_tool, Location))

# Compute summary statistics based on Chert_type and Strokes
stats_chert <- summaryBy(. ~ Chert_type + Strokes, data = STA_sel, FUN = nminmaxmeanmedsd)
stats_chert[1:3]

# Compute summary statistics based on Bamboo_sp and Strokes
stats_bamboo <- summaryBy(. ~ Bamboo_sp + Strokes, data = STA_sel, FUN = nminmaxmeanmedsd)
stats_bamboo[1:3]

# Compute summary statistics based on Chert_type, Bamboo_sp and Strokes
stats_chert_bamboo <- summaryBy(. ~ Chert_type + Bamboo_sp + Strokes, 
                                data = STA_sel, FUN = nminmaxmeanmedsd)
stats_chert_bamboo[1:4]
```


## Save as XLSX
```{r}
write_xlsx(list("Chert+Strokes" = stats_chert, "Bamboo+Strokes" = stats_bamboo, 
                "Chert+Bamboo+Strokes" = stats_chert_bamboo), 
           path = paste0(dir_stats, "/ISEA_use-wear_STA-stats.xlsx"))
```


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_3b_STA-stats")
knitr::kable(pkgs_cite)
```


## References

