---
title: "Plot tools' weights dataset for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_2c_Plot-Weight.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script imports and plots the tools' weights before and after the experiments.  
The script will:

1. Read in the original CSV-file  
2. Save data as Rbin and XLSX  
3. Plot 

```{r}
dir_in <- "analysis/raw_data/"
dir_data <- "analysis/derived_data/"
dir_plots <- "analysis/plots/"
```

Input CSV file file must be located in "`r paste0("./", dir_in)`".  
Plots will be saved in "`r paste0("./", dir_plots)`". Datasets will be saved in "`r paste0("./", dir_data)`". 

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(ggplot2)
library(ggrepel)
library(grateful)
library(knitr)
library(patchwork)
library(R.utils)
library(rmarkdown)
library(tidyverse)
library(writexl)
```


---


# Read in CSV file
```{r}
file_in <- list.files(dir_in, pattern = "weight\\.csv$", full.names = TRUE)
weights <- read.csv(file_in, check.names = FALSE)
str(weights)
head(weights)
```


---


# Save data
## As XLSX
```{r}
write_xlsx(weights, path = paste0(dir_data, "/ISEA_use-wear_weights.xlsx"))
```

## As Rbin
```{r}
saveObject(weights, file = paste0(dir_data, "/ISEA_use-wear_weights.Rbin"))
```

Rbin files (e.g. `ISEA_use-wear_weights.Rbin`) can be easily read into an R object (e.g. `rbin_data`) using the following code:
```{r, eval = FALSE}
library(R.utils)
rbin_data <- loadObject("ISEA_use-wear_weights.Rbin")
```


---


# Plot
## Pivot to long format for plotting
```{r}
weights_long <-  weights %>% 
                 pivot_longer(contains("Weight"), names_to = "State_full", values_to = "Weight [mg]") %>% 
                 mutate(State = factor(gsub("Weight_|_\\[mg\\]", "", State_full), 
                                       levels = c("before", "after")))
```

## Format column names for nice plotting
```{r}
color_name <- grep("type", names(weights_long), value = TRUE)
color_name_leg <- gsub("_", " ", color_name)
```

## Absolute weights
```{r}
             # Define plot
p_weights <- ggplot(weights_long, aes(x = State, y = .data[["Weight [mg]"]], color = .data[[color_name]])) +
             
             # Add points
             geom_point(size = 3) +
  
             # Add lines to connect points with identical"Sample ID
             geom_line(linewidth = 1, aes(group = Sample), show.legend = FALSE) +
  
             # Add Sample ID to points at "before" state
             geom_text_repel(aes(label = ifelse(State == "before", Sample, "")), show.legend = FALSE, 
                             hjust = 1, nudge_x = -0.1, direction = "y", 
                             min.segment.length = 0, segment.size = 0.3, seed = 123) +
   
             # Light theme
             theme_classic() +
  
             # Reduce margins around the plot
             scale_x_discrete(expand = expansion(add = c(0.9, 0.1))) +
  
             # The qualitative 'Set2' palette of RColorBrewer is colorblind friendly
             scale_color_brewer(palette = 'Set2') +
  
             # Remove xlab and use the clean name for the legend
             labs(x = NULL, color = color_name_leg)
```

## Weight differences
In order to improve the readability of the weight changes, only the differences are plotted.  

The difference is the weight after minus the weight before. Since all samples lost weight during the experiment, the absolute values are plotted in order to only show the magnitude of the difference.

```{r}
# Add column Weight_diff
weights <- mutate(weights, Weight_diff = abs(`Weight_after_[mg]` - `Weight_before_[mg]`))

          # Define plot
p_diff <- ggplot(weights, aes(x = .data[[color_name]], y = Weight_diff, color = .data[[color_name]])) +
          
          # Add points
          geom_point(size = 3) +
  
          # Add Sample ID to points at "before" state
          geom_text_repel(aes(label = Sample), show.legend = FALSE, hjust = 1, nudge_x = -0.1, 
                          direction = "y", min.segment.length = 0, segment.size = 0.3, seed = 123) +
  
          # Light theme
          theme_classic() +
  
          # Reduce margins around the plot
          scale_x_discrete(expand = expansion(add = c(0.7, 0.05))) +
  
          # The qualitative 'Set2' palette of RColorBrewer is colorblind friendly
          scale_color_brewer(palette = 'Set2') +
  
          # Remove xlab and use the clean name for the legend
          labs(x = NULL, y = "Weight difference [mg]",color = color_name_leg)
```

## Combine and plot plots
```{r}
p_all <- p_weights + p_diff + plot_layout(guides = 'collect')
plot(p_all)
```

## Save plot
```{r}
ggsave(plot = p_all, paste0(dir_plots, "/ISEA_use-wear_plot-weights.pdf"), width = 190, unit = "mm")
```


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_2c_Plot-Weight")
knitr::kable(pkgs_cite)
```


## References


