---
title: "Import dataset from Smarttester's sensors for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_1a_Import-Smarttester.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script imports and merges all single TXT-files (strokes + sensors) produced with the Inotec Smarttester during the experiments. The experiment involved 12 samples (2 types of flint and 2 types of bamboo, with 3 samples in each group) used for 2000 strokes.  
The script will:

1. Read in the original TXT-files   
2. Format and merge the data for each sample 
3. Combines the data from the 12 samples into one file  
4. Write an ODS file and save an R object ready for further analysis in R 

For each sample, each TXT file represents 10 back-and-fourth movements (= 20 strokes) for a given sensor. After 20 strokes, the bamboo worked piece was rotated.  
There were 5 positions (angles) for the bamboo: 0, -7.5, -15, +7.5 and +15°.  
There were 5 sensors: depth sensor (`Depth`), position (`X_position`) and velocity (`X_velocity`) along the X (movement) axis, and angle (`Angle`) and torque (`Torque`) of the rotary drive for the bamboo. Readings from each sensor are saved in a separate folder. It is possible to save up to 6 sensors' readings, so the folders are numbered from 1 to 6 (or multiples thereof, see below), even though we used only 5.  

In order to smooth the saving and storing of the files, the sensors' data were exported (written) after 20 strokes (hence each TXT file represents 20 strokes). However, the readings for each angle were saved in different folders, adding to the complexity of the folder structure.

There are 20 files per folder, where each file recorded 20 strokes, meaning that a total of 400 strokes were performed at each angle. With 5 angles, a total 2000 strokes was recorded per sample.  

With 20 files per folder and a folder for each combination of angle (5), sensor (5) and sample (12), a total of `r 20*5*5*12` TXT files were saved. Nevertheless, there were issues during the experiment with ISEA-EX3 so that the readings of the first 400 strokes with that sample were not usable.

Therefore, for each sample:

```{r, echo = FALSE, results = 'asis'}
# Create a sequence from 1 to 30 and omit every 6th entry, to mirror the numbering of the sensors by angle (1:5, 7:11...)
meas_index <- seq_len(30)[-seq(from = 6, to = 30, by = 6)]

# Split the sequence in groups of 5 to group the 5 sensor for each angle
angles <- split(meas_index, sort(meas_index %% 5))

# Name the list with the angles
names(angles) <- c("0°", "-7.5°", "-15°", "+7.5°", "+15°")

# Display the names of folders corresponding to each angle
for (i in 1:5) { 
  cat("- The folders '", paste0("Messung", angles[[i]][1], "..."), "' to '", paste0("Messung", angles[[i]][5], "..."), "' represent the readings for each of the 5 sensors,\n respectively, for Angle = ", names(angles)[i], "\n", sep = "")
}
```

In each folder, the files "SmartDatafiles00000000.txt" to "SmartDatafiles00000019.txt" did not record contiguous strokes but rather the strokes for a given sample and sensor at a given angle.  
For example:  

- in the folder "analysis/raw_data/Smarttester/ISEA-EX1/Messung1Date20231212131525", the file "SmartDatafiles00000000.txt" gives the readings of `Depth` at angle = 0° for sample ISEA-EX1 for the strokes 1-20  
- in the folder "analysis/raw_data/Smarttester/ISEA-EX1/Messung1Date20231212131525", the file "SmartDatafiles00000001.txt" gives the readings of `Depth` at angle = 0° for sample ISEA-EX1 for the strokes 101-120  
- in the folder "analysis/raw_data/Smarttester/ISEA-EX1/Messung1Date20231212131525", the file "SmartDatafiles00000019.txt" gives the readings of `Depth` at angle = 0° for sample ISEA-EX1 for the strokes 1901-1920  
- in the folder "analysis/raw_data/Smarttester/ISEA-EX1/Messung2Date20231212131525", the file "SmartDatafiles00000000.txt" gives the readings of `X_position` at angle = 0° for sample ISEA-EX1 for the strokes 1-20  
- in the folder "analysis/raw_data/Smarttester/ISEA-EX1/Messung7Date20231212131525", the file "SmartDatafiles00000000.txt" gives the readings of `Depth` at angle = -7.5° for sample ISEA-EX1 for the strokes 21-40 

```{r}
dir_in <- "D:/Data/ISEA_use-wear/3_Experiments_Inotec/Data"
dir_out <- "analysis/derived_data/"
```


Due to the huge number of TXT files (see below), it was not possible to upload them to GitHub. They were therefore stored and accessed locally in "`r dir_in`" for running the script. They can be accessed on Zenodo: *add DOI*

Formatted data will be saved in "`r dir_out`".

The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(grateful)
library(knitr)
library(R.utils)
library(readODS)
library(rmarkdown)
library(tidyverse)
```


---


# Read in TXT files

```{r}
# Create dataframe containing identification of folders
folders <- data.frame(bamboo = rep(c(0, -7.5, -15, 7.5, 15), each = 5), 
                      folder = seq_len(30)[-seq(from = 6, to = 30, by = 6)])

# Extract sample IDs from folder names
Samples_ID <- dir(dir_in)

# Create list, 1 element for each sample
Samples_data <- vector(mode = "list", length = length(Samples_ID)) 

# Create empty data.frame to contain a summary table of the number of detected TXT files per sample
num_txt <- data.frame(Sample = Samples_ID, NumberTXT = rep(NA, length(Samples_ID)))

# For each sample
for (i in seq_along(Samples_ID)) {
  
  # List all TXT files in dir_in/Samples_ID
  TXT_files <- list.files(paste(dir_in, Samples_ID[i], sep = "/"), 
                          pattern = "\\.txt$", recursive = TRUE, full.names = TRUE)
  
  # Display number of TXT files
  cat(length(TXT_files), "TXT files were found for sample", Samples_ID[i], "\n")
  
  # Save number of TXT files in num_txt
  num_txt[i, "NumberTXT"] <- length(TXT_files)
  
  # Import all TXT files 
  # Create list, 1 element for each TXT file
  TXT_data <- vector(mode = "list", length = length(TXT_files))

  # For each file
  for (j in seq_along(TXT_files)) {
    
  # Extract sensor's name and unit from the TXT file
               # Info on Line 3
  name_unit <- unlist(read.table(TXT_files[j], skip = 2, nrows = 1, sep = ";", 
                                 fileEncoding = 'WINDOWS-1252')) %>% 
                    
               # Format name and unit
               gsub("in ", "[", .)  %>% 
               paste0(., "]") %>%
               gsub("E2: ", "", .) %>% 
               gsub("Stroke", "X_position", .) %>% 
               gsub("Weg", "Depth", .) %>% 
               gsub("Velocity", "X_velocity", .)
  
  # Extract folder and file names from path
  messung <- as.numeric(gsub("Messung|Date[0-9]*", "", basename(dirname(TXT_files[j]))))
  smartfile <- as.numeric(gsub("SmartDatafiles000000|.txt", "", basename(TXT_files[j])))
  bamboo <- unique(folders[folders$folder == messung, "bamboo"])
  
  # Read in and format data
  TXT_data[[j]] <- read.table(TXT_files[j], skip = 4, sep = ";") %>%
                   mutate(Sample = Samples_ID[i], File = smartfile, Bamboo_position = bamboo, 
                          Sensor_name = name_unit) %>%
                   rename(Sensor_value = "V1", Step = "V2")
  }
  
  # rbind all files per sample
  Samples_data[[i]] <- do.call(rbind, TXT_data)
  str(Samples_data[[i]])
  head(Samples_data[[i]])
}

# rbind all files for all samples
all_data <- do.call(rbind, Samples_data) %>% 
            arrange(Sample, File, Bamboo_position, Sensor_name, Step)
str(all_data)
head(all_data)
```

The number of TXT files read in for each sample is as follow: 

```{r, echo = FALSE}
num_txt_order <- num_txt %>% 
                 mutate(SampleNr = as.numeric(gsub("ISEA-EX", "", num_txt$Sample))) %>% 
                 arrange(SampleNr)
num_txt_order[1:2]
```

In total, `r sum(num_txt_order$NumberTXT)` TXT files have been read in.


---


# Format data

## Pivot to wider format
Create one column (variable) for each sensor and create an empty column for the stroke number.

```{r}
wide_data <- pivot_wider(all_data, names_from = Sensor_name, values_from = Sensor_value) %>% 
             mutate(Bamboo_position = factor(Bamboo_position, levels = unique(folders$bamboo)), 
                    StrokeNr = NA) %>% 
             arrange(Sample, File, Bamboo_position, Step) %>% 
             as.data.frame()
str(wide_data)
head(wide_data)
```


## Add stroke number
```{r}
# Split by sample to work on each sample separately
split_sample <- split(wide_data, wide_data$Sample)

# Create empty data.frame to contain a summary table of the number of detected strokes per sample
num_stroke <- data.frame(Sample = names(split_sample), NumberStrokes = rep(NA, length(names(split_sample))))

# For each sample
for (i in seq_along(split_sample)) {
  
  # Start the first stroke at the first row
  # For ISEA-EX3, the first stroke in the dataset is stroke number 401
  if (grepl("ISEA-EX3", names(split_sample)[i])) {
    split_sample[[i]][1, "StrokeNr"] <- 401
  } else {
    split_sample[[i]][1, "StrokeNr"] <- 1
  }
  
  # Calculate the lagged difference between successive values of the X position ("Stroke [mm]")
  diff_stroke <- diff(split_sample[[i]][["X_position [mm]"]])
  
  # Duplicate the first value to have the same length as the original values
  diff_stroke <- c(diff_stroke[1], diff_stroke)
  
  # Identify differences of 0
  diff_stroke_0 <- which(diff_stroke == 0)
  
  # Replace the 0s with the previous value
  diff_stroke[diff_stroke_0] <- diff_stroke[diff_stroke_0 - 1]
  
  # Create empty vector to contain the result
  sign_diff_stroke <- vector(mode = "logical", length = length(diff_stroke))
  
  # For each difference (starting at the second value)
  for (j in 2:length(diff_stroke)) {
    
    # Identify inversions in the sign of the difference, i.e. when the drive starts moving in the other direction
    sign_diff_stroke[j] <- sign(diff_stroke[j]) != sign(diff_stroke[j-1])
  }
  
  # Identify inversions in the sign of the differences
  inv_stroke <- which(sign_diff_stroke)
  
  # Fill the data.frame num_stroke with the number of inversions + 1 (because it starts at stroke number 2)
  num_stroke[i, "NumberStrokes"] <- length(inv_stroke) + 1
  
  # For each inversion in the sign of the difference, 
  # using the index of the inversion - 1 in order to compensate for the lag,
  # increment the stroke number.
  # Stroke number starts at 2 because the 1st stroke was added manually at the 1st row
  # In case of ISEA-EX3, the stroke number starts at 402 because the first 400 strokes were not recorded properly
  if (grepl("ISEA-EX3", names(split_sample)[i])) {
    split_sample[[i]][inv_stroke - 1, "StrokeNr"] <- seq_along(inv_stroke) + 401
  } else {
    split_sample[[i]][inv_stroke - 1, "StrokeNr"] <- seq_along(inv_stroke) + 1
  }
  
  # Fill the stroke number down to fill in the rows of identical stroke numbers
  split_sample[[i]] <- fill(split_sample[[i]], StrokeNr)
}

# rbind all samples into a single data.frame
final_data <- do.call(rbind, split_sample)
```

The number of inversions (= strokes) detected in the sensor's readings for each sample is as follow: 

```{r, echo = FALSE}
num_stroke_order <- num_stroke %>% 
                    mutate(SampleNr = as.numeric(gsub("ISEA-EX", "", num_stroke$Sample))) %>% 
                    arrange(SampleNr)
num_stroke_order[1:2]
```


## Reorder columns and renumber rows
```{r}
final_data <- final_data[c(2:4, 1, 10, 5:9)]
row.names(final_data) <- NULL
```


## Check result
```{r}
str(final_data)
head(final_data)
```


---


# Save data
```{r}
write_ods(final_data, path = paste0(dir_out, "/ISEA_use-wear_Smarttester.ods"))
```


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_1a_Import-Smarttester")
knitr::kable(pkgs_cite)
```


## References


