---
title: "Plots for the ISEA use-wear dataset"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 4
    html_preview: false
bibliography: ISEA_3c_STA-plots.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
The script plots all STA variables for the ISEA use-wear dataset.   

```{r}
dir_in  <- "analysis/derived_data"
dir_plots <- "analysis/plots"
```

Input Rbin data file must be located in "`r paste0("./", dir_in)`".  
Plots will be saved in "`r paste0("./", dir_plots)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(doBy)
library(factoextra)
library(ggplot2)
library(grateful)
library(gridExtra)
library(knitr)
library(R.utils)
library(RColorBrewer)
library(rmarkdown)
library(tidyverse)
```


---


# Read in and format data
## Read in Rbin file
```{r}
STA <- list.files(dir_in, pattern = "STA\\.Rbin$", full.names = TRUE) %>% 
       loadObject()
str(STA)
head(STA)
```


## Calculate the mean per sample
There are `r length(unique(STA$Sample))` samples, with height maps acquired on `r length(unique(STA$Side))` sides (`r paste(unique(STA$Side), collapse = " and ")`), 2 locations per side, and at `r paste(levels(STA$Strokes), collapse = " and ")` strokes. This equals to `r nrow(STA)` data points.

In order to avoid cluttering the plots, the mean per sample (n = `r length(unique(STA$Sample))` at `r paste(levels(STA$Strokes), collapse = " and ")` strokes) is plotted rather than every single data point.

```{r}
# Exclude Chert_tool, Location and NMP columns from data because they are numeric yet not relevant for plots
STA_sel <- select(STA, !c(Chert_tool, Location, NMP))

# Calculate mean based on Sample + Strokes
# Bamboo_sp and Chert_type are listed as factors in order to keep these columns
# keep.names = TRUE is important for the matching on names in the next steps
STA_mean <- summaryBy(.~ Sample + Bamboo_sp + Chert_type + Strokes, data = STA_sel, FUN = mean, keep.names = TRUE)
```


## Add units to headers for plotting
```{r}
# Get units from comment(STA)
table_units <- comment(STA) %>% 
               data.frame(Parameter = names(.), Unit = ., row.names = NULL) %>% 
  
               # Exclude NMP because it won't be plotted
               filter(Parameter != "NMP") %>% 
  
               # Paste parameter name and unit together in a new column
               mutate(Param_unit = paste0(Parameter, " [", Unit, "]")) 

# Remove > and < symbols
table_units$Param_unit <- gsub(">|<", "", table_units$Param_unit)

# Adjust column names
colnames(STA_mean)[colnames(STA_mean) %in% table_units$Parameter] <- table_units$Param_unit
```


## Define parameter types for grouping plots
```{r}
table_units[table_units$Parameter %in% c("Sq", "Ssk", "Sku", "Sp", "Sv", "Sz", "Sa"), 
            "Type"] <- "ISO height"

table_units[grepl("^V[a-z]{1,2}", table_units$Parameter), "Type"] <- "ISO volume"

table_units[table_units$Parameter %in% c("Smr", "Smc", "Sxp", "Sal", "Str", "Std", "Ssw", "Sdq", "Sdr"), 
            "Type"] <- "ISO other"

table_units[grepl("furrows|direction|isotropy", table_units$Parameter), 
            "Type"] <- "Furrow, direction and isotropy"

table_units[grepl("Asfc|(?i)eplsar|Smfc", table_units$Parameter), "Type"] <- "SSFA"

table_units[c("Param_unit", "Type")]
```


---


# Plot each surface parameter in a boxplot
## Define variables
Here we define which columns are used for the plots.  

```{r}
# Columns to be used to group on the x-axis
x_var <- "Strokes"

# colors
grp_colors <- "Chert_type"
color_name_leg <- gsub("_", " ", grp_colors)

# shapes
grp_shapes <- "Bamboo_sp"
shape_name_leg <- gsub("_", " ", grp_shapes)
```


## Create lists to receive the plots
```{r}
p_box <- vector(mode = "list", length = length(unique(table_units$Type)))
names(p_box) <- unique(table_units$Type)
```


## Plots 
```{r}
# Plot for every set of parameters
for (i in names(p_box)) {
  
  # Subset parameter names by type
  y_var_i <- table_units[table_units$Type == i, "Param_unit"]
  
  # Select columns that will be used in the plotting
  data_i <- select(STA_mean, all_of(c("Sample", "Bamboo_sp", x_var, y_var_i, grp_colors, grp_shapes))) %>%
  
            # Pivot to longer format for facet plots
            pivot_longer(all_of(y_var_i), names_to = "Parameter", values_to = "Value")
  
  # Plot        # Define aesthetics
  p_box[[i]] <- ggplot(data_i, aes(x = .data[[x_var]], y = Value,
                                   color = .data[[grp_colors]], shape = .data[[grp_shapes]])) +

                # Add points
                geom_point(size = 2) +
    
                # Add lines to connect points with identical"Sample ID
                geom_line(linewidth = 0.5, aes(group = Sample), 
                          show.legend = FALSE) +
    
                # Facet plot by 'Parameter' (= 'y_var_i')
                facet_wrap(~ Parameter, scales = "free_y") +
    
                # Light theme
                theme_classic() +
    
                # Italicize species names in legend
                guides(shape = guide_legend(theme = theme(legend.text = element_text(face = "italic")))) +
    
                # Reduce font size of the facet labels so that they fit in the box
                theme(strip.text = element_text(size = 6)) +
  
                # The qualitative 'Set2' palette of RColorBrewer is colorblind friendly
                scale_color_brewer(palette = 'Set2') +
  
                # Remove xlab and use the clean name for the legend
                labs(y = NULL, title = paste0(i, " (mean per sample)"), 
                     color = color_name_leg, shape = shape_name_leg)
}

# Print plot
print(p_box)

# Save list of plots
ggsave(filename = "ISEA_use-wear_STA-plots.pdf", 
       path = dir_plots, width = 190, units = "mm", 
       plot = marrangeGrob(p_box, nrow = 1, ncol = 1, top = NULL))
```


---


# PCA
TO DO


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_3c_STA-plots")
knitr::kable(pkgs_cite)
```


## References

