---
title: "Import dataset from the surface texture analysis for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_2a_STA-import.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---


```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---


# Goal of the script
This script formats the output of the resulting files from applying surface texture analysis to a sample of experimental lithics.
The script will:

1. Read in the original files  
2. Format the data  
3. Write XLSX file and save R objects ready for further analysis in R  

```{r}
dir_in  <- "analysis/raw_data"
dir_out <- "analysis/derived_data"
```

Raw data must be located in "`r paste0("./", dir_in)`".  
Formatted data will be saved in "`r paste0("./", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(grateful)
library(knitr)
library(R.utils)
library(rmarkdown)
library(tidyverse)
library(writexl)
```


---


# Read and format data
## Read in CSV file
```{r}
# Loop through list of CSV files and read in
STA <- list.files(dir_in, pattern = "STA\\.csv$", full.names = TRUE) %>% 
       read.csv(header = FALSE, na.strings = "*****", fileEncoding = 'WINDOWS-1252')
```


## Select relevant columns and rows
```{r}
STA_keep_col  <- c(3, 11:46)                               # Define columns to keep
STA_keep_rows <- which(STA[[1]] != "#")                    # Define rows to keep
STA_keep      <- STA[STA_keep_rows, STA_keep_col] # Subset rows and columns
```


## Identify results using frame numbers
```{r}
frames <- as.numeric(unlist(STA[1, STA_keep_col]))
ID <- which(frames == 26)
ISO <- which(frames == 28)
furrow <- which(frames == 29)
diriso <- which(frames %in% 30:31)
SSFA <- which(frames %in% 32:33)
```


## Add headers
```{r}
# Get headers from 2nd row
colnames(STA_keep) <- STA[2, STA_keep_col] %>% 
  
                      # Convert to valid names
                      make.names() %>% 
  
                      # Delete repeated periods
                      gsub("\\.+", "\\.", x = .) %>% 
  
                      # Delete periods at the end of the names
                      gsub("\\.$", "", x = .)
  
# Keep parameter name before the first period for ISO
colnames(STA_keep)[ISO] <- strsplit(names(STA_keep)[ISO], ".", fixed = TRUE) %>% 
                           sapply(`[[`, 1)

# Keep parameter name after the last period for SSFA
colnames(STA_keep)[SSFA] <- gsub("^([A-Za-z0-9]+\\.)+", "", colnames(STA_keep)[SSFA])

# Edit header non-measured point ratios
colnames(STA_keep)[ID] <- "NMP"

# Edit header for name of surfaces
colnames(STA_keep)[1] <- "Name"

# Check results
str(STA_keep)
head(STA_keep)
```


## Extract units
```{r}
# Filter out rows which contains units
n_units <- filter(STA, V4 == "<no unit>") %>% 
          
           # Keep only unique/distinct rows of units
           distinct() %>% 
  
           # Number of unique/distinct rows of units
           nrow()

if (n_units != 1) {
  stop("The different datasets have different units")
} else {
  # Extract unit line from 3rd row for considered columns
  STA_units <- unlist(STA[3, STA_keep_col[-1]])

  # Get names associated to the units
  names(STA_units) <- colnames(STA_keep)[-1]

  # Combine into a data.frame for export
  units_table <- data.frame(variable = names(STA_units), units = STA_units)
  row.names(units_table) <- NULL
}

# Check results
units_table
```


## Split column 'Name' 
```{r}
STA_keep[c("Specimen", "Objective", "Side", "Location", "Strokes")] <- STA_keep$Name %>% 
  
                                                                       # Split by " --- " and keep 2nd element
                                                                       str_split_i(" --- ", i = 2) %>% 
                                                                       
                                                                       # Delete ".mnt"
                                                                       gsub(".mnt", "", .) %>% 
   
                                                                       # Split by "_" into a matrix with 7 cols
                                                                       str_split_fixed("_", n = 7) %>% 
  
                                                                       # Keep all cols but cols 2 and 6
                                                                       .[, -c(2, 6)]
```


## Split column 'Specimen' 
```{r}
STA_keep[c("Chert_type", "Chert_tool")] <- STA_keep$Specimen %>% 
                                           
                                           # Split by "-" and keep 3rd element   
                                           str_split_i("-", i = 3) %>% 
  
                                           # Split after the first capital letter
                                           str_split("(?<=[A-Z]{1})", simplify = TRUE)
```


## Add column with sample ID
```{r}
# Create a data.frame with correspondence between chert type, chert tool and sample ID
sample_ID <- data.frame(Sample = paste0("ISEA-EX", 1:12), 
                        Chert_type = rep(c("A", "B"), times = 6),
                        Chert_tool = rep(1:6, each = 2))

# Merge the data.frames by chert type and chert tool
STA_keep_ID <- merge(STA_keep, sample_ID, by = c("Chert_type", "Chert_tool"))
```


## Keep only numerical parts in columns "Location" and "Strokes"
```{r}
STA_keep_ID$Location <- gsub("loc", "", STA_keep_ID$Location)
STA_keep_ID$Strokes <- factor(gsub("strokes", "", STA_keep_ID$Strokes))
```


## Convert all parameter variables to numeric
```{r}
STA_keep_ID <- type_convert(STA_keep_ID)
```


## NMP ratios
The summary statistics for NMP ratios are:  
```{r, echo = FALSE}
summary(STA_keep_ID[["NMP"]])
```

All seems good.


## Re-order columns and add units as comment 
```{r}
STA_final <- select(STA_keep_ID, Sample, Chert_type, Chert_tool, Objective:Strokes, NMP:NewEplsar)
comment(STA_final) <- STA_units
```

Type `comment(STA_final)` to check the units of the parameters.


## Check the result
```{r}
str(STA_final)
head(STA_final)
```


---


# Save data
## As XLSX
```{r}
write_xlsx(list("data" = STA_final, "units" = units_table), path = paste0(dir_out, "/ISEA_use-wear_STA.xlsx"))
```


## As Rbin
```{r}
saveObject(STA_final, file = paste0(dir_out, "/ISEA_use-wear_STA.Rbin"))
```

Rbin files (e.g. `ISEA_use-wear_STA.Rbin`) can be easily read into an R object (e.g. `rbin_data`) using the following code:
```{r, eval = FALSE}
library(R.utils)
rbin_data <- loadObject("ISEA_use-wear_STA.Rbin")
```


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_2a_STA-import")
knitr::kable(pkgs_cite)
```


## References

