---
title: "Plot tools' weights dataset for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_2c_Plot-Weight.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script imports and plots the tools' weights before and after the experiments.  
The script will:

1. Read in the original CSV-file  
2. Save data as Rbin and XLSX  
3. Plot 

```{r}
dir_in <- "analysis/raw_data/"
dir_out <- "analysis/plots/"
```

Input CSV file file must be located in "`r paste0("./", dir_in)`".  
Plots will be saved in "`r paste0("./", dir_out)`".

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(ggplot2)
library(grateful)
library(knitr)
library(R.utils)
library(rmarkdown)
library(tidyverse)
library(writexl)
```


---


# Read in CSV file
```{r}
file_in <- list.files(dir_in, pattern = "weight\\.csv$", full.names = TRUE)
weights <- read.csv(file_in, check.names = FALSE)
str(weights)
head(weights)
```


---


# Save data
## As XLSX
```{r}
write_xlsx(weights, path = paste0(dir_out, "/ISEA_use-wear_weights.xlsx"))
```

## As Rbin
```{r}
saveObject(weights, file = paste0(dir_out, "/ISEA_use-wear_weights.Rbin"))
```

Rbin files (e.g. `ISEA_use-wear_weights.Rbin`) can be easily read into an R object (e.g. `rbin_data`) using the following code:
```{r, eval = FALSE}
library(R.utils)
rbin_data <- loadObject("ISEA_use-wear_weights.Rbin")
```


---


# Plot


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_2c_Plot-Weight")
knitr::kable(pkgs_cite)
```


## References


