---
title: "Plot Smarttester's sensor data for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_1b_Smarttester-plot.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script plots the sensors data from the Smarttester.  


```{r}
dir_in <- "analysis/derived_data/"
dir_plots <- "analysis/plots/"
```

Input CSV file file must be located in "`r paste0("./", dir_in)`".  
Plots will be saved in "`r paste0("./", dir_plots)`". 

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(ggplot2)
library(grateful)
library(gridExtra)
library(knitr)
library(R.utils)
library(rmarkdown)
library(tidyverse)
library(viridis)
```


---


# Read in Rbin file
```{r}
ST_file <- list.files(dir_in, pattern = "Smarttester\\.Rbin$", full.names = TRUE)
ST_data <- loadObject(ST_file)
str(ST_data)
head(ST_data)
```


---


# Plot

## Define variables
Define dependent (numeric) variables to plot on the Y axis

```{r}
# Y-axis
ST_y <- c("X_velocity [mm/s]", "Depth [mm]", "Angle [°]", "Torque [Nm]")

# X-axis
X_name <- grep("X_position", names(ST_data), value = TRUE)
X_name_leg <- gsub("_", " ", X_name)
```


## Subset and prepare data for plots
For readability, only every 20th stroke will be plotted. The first 50 strokes going along the positive X axis (i.e. odd strokes from 1 to 100) will also be plotted separately.

### Every 20th strokes
```{r}
# Create a sequence of every 20th strokes, including first and before last strokes
# Before last is necessary because of bidirectional strokes:
# the odd strokes go in one X direction, while the even strokes go in the other
stroke_20 <- seq(1, 2000, by = 20) %>% 
             c(max(ST_data$StrokeNr) - 1) 

# Extract data for these strokes
data_every20 <- filter(ST_data, StrokeNr %in% stroke_20) %>% 

                # Pivot to longer format for facet plots
                pivot_longer(all_of(ST_y), names_to = "parameter", values_to = "value") %>% 
  
                # Convert to data.frame
                as.data.frame()

# Check result
str(data_every20)
head(data_every20)
```

### First 50 positive (odd) strokes
```{r}
# Extract data for the first 50 odd strokes
data_1to50 <- filter(ST_data, StrokeNr %in% seq(1, 100, 2)) %>% 

              # Pivot to longer format for facet plots
              pivot_longer(all_of(ST_y), names_to = "parameter", values_to = "value") %>% 
  
              # Convert to data.frame
              as.data.frame()

# Check result
str(data_1to50)
head(data_1to50)
```


## Prepare receiving lists
Data for the first 400 strokes of sample ISEA-EX3 were not recorded, so the first 50 positive strokes cannot be plotted for that sample.

```{r}
# Sample IDs
sample_ID <- levels(factor(ST_data$Sample)) %>% 
             data.frame(Sample = ., SampleNr = .) %>% 
             mutate(SampleNr = as.numeric(gsub("ISEA-EX", "", SampleNr))) %>% 
             arrange(SampleNr)

# Create lists to receive the plots
p_every20 <- p_1to50 <- vector(mode = "list", length = nrow(sample_ID))
names(p_every20) <- names(p_1to50) <- sample_ID$Sample

# Delete ISEA-EX3 from p_1to50
p_1to50[["ISEA-EX3"]] <- NULL
```


## Plots
### Every 20th stroke
```{r}
# Plot for every sample
for (i in names(p_every20)) {
  
  # Filter data per sample
  data_every20_i <- filter(data_every20, Sample == i)
  
                    # Prepare plot
  p_every20[[i]] <- ggplot(data = data_every20_i, 
                           aes(x = .data[[X_name]], y = value, color = StrokeNr)) + 
  
                    # Draw one line per stroke
                    geom_line(aes(group = StrokeNr), alpha = 0.7, linewidth = 0.4) +
  
                    # Facet plot by 'parameter' (= 'ST_y')
                    facet_wrap(~ parameter, scales = "free_y") +
  
                    # The palette 'viridis' is printer- and colorblind-friendly, and perceptually uniform
                    scale_color_viridis(trans = "reverse") + 
  
                    # Light theme, and remove x-axis text and ticks
                    theme_classic() + 
                    
                    # No y-axis label and title from sample_ID$Sample
                    labs(y = NULL, title = i, x = X_name_leg)
}

# Print list of plots
print(p_every20)

# Save list of plots
ggsave(filename = "ISEA_use-wear_plot-Smarttester_all-strokes.pdf", 
       path = dir_plots, width = 190, units = "mm", 
       plot = marrangeGrob(p_every20, nrow = 1, ncol = 1, top = NULL))
```

### First 50 strokes
Data for the first 400 strokes of sample ISEA-EX3 were not recorded, so the first 50 positive strokes cannot be plotted for that sample.

```{r}
# Plot for every sample
for (i in names(p_1to50)) {
  
  # Filter data per sample
  data_1to50_i <- filter(data_1to50, Sample == i)
  
                  # Prepare plot
  p_1to50[[i]] <- ggplot(data = data_1to50_i, 
                         aes(x = .data[[X_name]], y = value, color = StrokeNr)) + 
  
                  # Draw one line per stroke
                  geom_line(aes(group = StrokeNr), alpha = 0.7, linewidth = 0.4) +
  
                  # Facet plot by 'parameter' (= 'ST_y')
                  facet_wrap(~ parameter, scales = "free_y") +
  
                  # The palette 'viridis' is printer- and colorblind-friendly, and perceptually uniform
                  scale_color_viridis(trans = "reverse") + 
  
                  # Light theme, and remove x-axis text and ticks
                  theme_classic() + 
                    
                  # No y-axis label and title from sample_ID$Sample
                  labs(y = NULL, title = i, x = X_name_leg)
}

# Print list of plots
print(p_1to50)

# Save list of plots
ggsave(filename = "ISEA_use-wear_plot-Smarttester_50strokes.pdf", 
       path = dir_plots, width = 190, units = "mm", 
       plot = marrangeGrob(p_1to50, nrow = 1, ncol = 1, top = NULL))
```


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_1b_Smarttester-plot")
knitr::kable(pkgs_cite)
```


## References


