---
title: "Plot Smarttester's sensor data for the ISEA use-wear project"
author: "Ivan Calandra"
date: "`r format(Sys.time(), usetz = TRUE)`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
bibliography: ISEA_2a_Plot-Smarttester.bib
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include = FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)
```


---

# Goal of the script
This script plots the sensors data from the Smarttester.  


```{r}
dir_in <- "analysis/derived_data/"
dir_plots <- "analysis/plots/"
```

Input CSV file file must be located in "`r paste0("./", dir_in)`".  
Plots will be saved in "`r paste0("./", dir_plots)`". 

The knit directory for this script is the project directory.


---


# Load packages
```{r}
library(ggplot2)
library(grateful)
library(knitr)
library(R.utils)
library(rmarkdown)
library(tidyverse)
```


---


# Read in CSV file
```{r}
ST_file <- list.files(dir_in, pattern = "Smarttester\\.Rbin$", full.names = TRUE)
ST_data <- loadObject(ST_file)
str(ST_data)
head(ST_data)
```


---


# Plot

## Define variables
Define dependent (numeric) variables to plot on the Y axis

```{r}
ST_y <- c("X_velocity [mm/s]", "Depth [mm]", "Angle [°]", "Torque [Nm]")
```


## Subset and prepare data for plots
For readability, only every 20th stroke will be plotted. The first 50 strokes will also be plotted separately.

### Every 20th strokes
```{r}
# Create a sequence of every 20th strokes, including first and last strokes
stroke_20 <- seq(1, 2000, by = 20) %>% 
             c(max(ST_data$StrokeNr)) 

# Extract data for these strokes
data_every20 <- filter(ST_data, StrokeNr %in% stroke_20) %>% 

                # Pivot to longer format for facet plots
                pivot_longer(all_of(ST_y), names_to = "parameter", values_to = "value") %>% 
  
                # Convert to data.frame
                as.data.frame()

# Check result
str(data_every20)
head(data_every20)
```

### First 50 strokes
```{r}
# Extract data for the first 50 strokes
data_1to50 <- filter(ST_data, StrokeNr %in% 1:50) %>% 

              # Pivot to longer format for facet plots
              pivot_longer(all_of(ST_y), names_to = "parameter", values_to = "value") %>% 
  
              # Convert to data.frame
              as.data.frame()

# Check result
str(data_1to50)
head(data_1to50)
```


## Plots
```{r}

```


---


# sessionInfo()

```{r}
sessionInfo()
```


---


# Cite R packages used

```{r, echo = FALSE}
pkgs_cite <- cite_packages(pkgs = "Session", omit = NULL, output = "table", include.RStudio = TRUE, 
                           out.dir = "analysis/scripts", bib.file = "ISEA_2a_Plot-Smarttester")
knitr::kable(pkgs_cite)
```


## References


